---
title: "PA2_template.Rmd"
author: "Carlo Bonini"
date: "Sunday, July 20, 2014"
output: html_document
---


Here is the preprocessing code:
```{r}

raw<-read.csv("activity.csv")

good<-complete.cases(raw)

goodies<-raw[good,]

```



Here I make a simple histogram of steps taken  on a daily basis: 

```{r, echo=TRUE}
library(sqldf)

day_steps<-sqldf("select distinct(date), sum(steps) from goodies group by date")

hist(day_steps[,2],xlab="steps taken each day",main="Histogram of steps taken each day")

```

While in this other code chunk I calculate both Mean and Median:


```{r, echo=TRUE}

average_daily_steps<-mean(day_steps[,2])

Median_daily_steps<-median(day_steps[,2])

average_daily_steps

Median_daily_steps
```


First, I compute the raw data in order to have a Tiny data sets to further computations,and then I make a simple time_series plot using the Base-Plotting system of the average steps taken by interval, averaged across all days: 

```{r, echo=TRUE}
average_daily_steps_per_interval<-sqldf("select interval, avg(steps) from goodies group by interval")

plot(average_daily_steps_per_interval[,1],
     average_daily_steps_per_interval[,2],type="l",main="Average Steps per interval",xlab="Intervals",ylab="average daily steps")

```

Finally I calculate the interval during the day that scored the highest number of steps (the result also the number of steps) :

```{r, echo=TRUE}

max_steps_interval<-subset(average_daily_steps_per_interval,
                           average_daily_steps_per_interval[,2]==
                               max(average_daily_steps_per_interval[,2]))

max_steps_interval
```

The total number of NAs is 
```{r, echo=TRUE}

Number_of_NA<-as.data.frame(length(raw[is.na(raw)]))

Number_of_NA
```

And the way I intended to substitute them is by subsitute each of them with the average of the interval calculated above(across all days):
The solution I decided to use to accomplish this is quite a bit complicated(I know there are simpler way to do that);
- First, I merged the data frame calculated early of the average steps taken by interval with the raw data, so that for each row we have the mean of the relative interval
- Then I subsitute the NA's with the relative mean that is in column 4 after the merge.

```{r, echo=TRUE}

merged_raw_average<-merge(average_daily_steps_per_interval,raw,by="interval")


for (i in 1:nrow(merged_raw_average)) {

  if (is.na(merged_raw_average[i,3])){
  
  merged_raw_average[i,3]<-merged_raw_average[i,2]

  }

}

```


I made a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

The  values don't differ from the estimates from the first part of the assignment.

```{r, echo=TRUE}

day_steps_new<-sqldf("select distinct(date), sum(steps) from merged_raw_average group by date")

hist(as.numeric(as.character(day_steps_new[,2])),xlab="steps taken each day",main="Histogram of steps taken each day")

average_daily_steps_new<-mean(day_steps_new[,2])

median_daily_steps_new<-median(day_steps_new[,2],na.rm = FALSE)

average_daily_steps_new

median_daily_steps_new

```


So, the new data frame has 4 column and no NA's.
I transform the column of the date into a more useful format by assigning the weekdays value to each cell, so that is easy to understand whether a day is weekend or weekday.


```{r, echo=TRUE}

merged_raw_average[,4]<-as.Date(merged_raw_average[,4],origin="1960-01-01")

merged_raw_average[,4]<-weekdays(merged_raw_average[,4])

```

Here I have subset the data frame based on the fact that in column 3 there were weekend days (in italian: sabato and domenica) and subset again based on all the other values.
Then i bind them together again after assigning a factor variable that indicates whether is weekday or weekend:

In the end I subset the final database using the package sqldf in order to get the average per interval across all days as above; I get two different dataframes per weekends data and weekdays data.


```{r, echo=TRUE}

weekend<-subset(merged_raw_average,merged_raw_average$date=="sabato" | merged_raw_average$date=="domenica")

weekend["type_of_day"]<-"weekend"

weekday<-subset(merged_raw_average,merged_raw_average$date!="sabato"&merged_raw_average$date!="domenica")

weekday["type_of_day"]<-"weekday"

final_database<-rbind(weekday,weekend)

weekday_average_daily_steps<-sqldf("select interval, avg(steps) from final_database where type_of_day= 'weekday' group by interval")

weekend_average_daily_steps<-sqldf("select interval, avg(steps) from final_database where type_of_day= 'weekend' group by interval")


```

Finally, we can make the plot using the base plotting system:

```{r, echo=TRUE}

plot(weekday_average_daily_steps[,1],
     weekday_average_daily_steps[,2],type="l",main="WEEKDAY",xlab="Intervals",ylab="average daily steps")

plot(weekend_average_daily_steps[,1],
     weekend_average_daily_steps[,2],type="l",main="WEEKEND",xlab="Intervals",ylab="average daily steps")



```
